A Sound and Complete Deep Embedding of Arbitrary Normal Modal Logics in Coq

### 2025-08-20 paper/COLA/paper.tex
Some rephrasing

\paragraph{The completeness proof} We almost have all the pieces to prove the completeness theorem. The next lemma
shows that a formula holds in the canonical model if, and only if, it is provable in the deductive system.
\begin{minted}{coq}
Lemma determination: forall p, (M |= p) <-> (A; Empty |-- p).
\end{minted}
The proof is divided in two more lemmas for each implication direction using the truth lemma. Finally, the desired completeness theorem is stated simply as follows:
\begin{minted}{coq}
Theorem completeness: forall p, (Empty ||= p) -> (A; Empty |-- p).
Proof.
  (* Comments present in the source code omitted. *)
  intros p.
  apply contrapositive; intros.
  - apply classic.
  - assert (Consistent A Empty).
    + now apply nonderivation_implies_consistency with p.
    + assert ((M |= p) -> False); intros.
      * apply determination with p in H2.
        contradiction.
      * apply H2, H0; easy.
Qed.
\end{minted}
The proof proceeds by the contrapositive.
Suppose that \mintinline{coq}{A; Empty |-- p} does not hold.
As this system can't derive every formula, it has to be consistent, which we can safely assume.
Since the empty context is clearly consistent, it
can be extended into the canonical model \verb|M|. By the fact that
\mintinline{coq}{A; Empty |-- p} is false, by the determination lemma, we know
that \verb|p| is not true on the canonical model \verb|M|. Conclusion then follows directly
by our current hypotheses, namely that \mintinline{coq}{Empty ||= p} and \mintinline{coq}{(A; Empty |-- p) -> False} (which includes showing that all formulas in the empty set are valid).

>>>>

\paragraph{The completeness proof} We almost have all the pieces to prove the completeness theorem. The next lemma
shows that a formula holds in the canonical model if, and only if, it is provable in the deductive system.
\begin{minted}{coq}
Lemma determination: forall p, (M |= p) <-> (A; Empty |-- p).
\end{minted}
The proof is divided in two more lemmas for each implication direction using the truth lemma. Finally, the desired completeness theorem is stated simply as follows:
\begin{minted}{coq}
Theorem completeness: forall p, (Empty ||= p) -> (A; Empty |-- p).
Proof.
  (* Comments present in the source code omitted. *)
  intros p.
  apply contrapositive; intros.
  - apply classic.
  - assert (Consistent A Empty).
    + now apply nonderivation_implies_consistency with p.
    + assert ((M |= p) -> False); intros.
      * apply determination with p in H2.
        contradiction.
      * apply H2, H0; easy.
Qed.
\end{minted}
The proof proceeds by the contrapositive.
Suppose that \mintinline{coq}{A; Empty |-- p} does not hold.
As this system can't derive every formula, it has to be consistent, which we can safely assume.
Since the empty context is clearly consistent, it
can be extended into the canonical model \verb|M|. By the fact that
\mintinline{coq}{A; Empty |-- p} is false, by the determination lemma, we know
that \verb|p| is not true on the canonical model \verb|M|. Conclusion then follows directly
by our hypotheses, \mintinline{coq}{Empty ||= p} and \mintinline{coq}{(A; Empty |-- p) -> False} (which includes showing that all formulas in the empty set are valid).
